---
title: "Results"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r}
# load packages and data
library(tidyverse)
source("https://raw.githubusercontent.com/HenrikEckermann/in_use/master/reporting.R")



data(Howell1, package = "rethinking")
df <- Howell1

df <- mutate(df, gender = ifelse(male == 1, "male", ifelse(male == 0, "female", NA))) %>%
      mutate(gender = as.factor(gender)) %>%
      select(-male)

```





```{r}
# descriptive statistics
sum_funs <- c("mean", "sd", "min", "max")
sum_rows <- map_dfr(sum_funs, function(fun) {
  df_cont <- df %>%
    summarise_if(is.numeric, fun) %>%
    mutate(statistic = fun) %>%
    mutate_if(is.numeric, round, 2) %>%
    select(statistic, height, weight, age)
})
table_1 <- sum_rows %>% 
  pivot_longer(-statistic, names_to = "variable", values_to = "value") %>%
  pivot_wider(names_from = statistic, values_from = value)
```



```{r}
# main analysis
model <- lm(height ~ weight + age + gender, data = df)
coefs <- model %>%
  summary() %>%
  coefficients %>%
  as.data.frame() %>%
  rownames_to_column("Parameter") %>%
  mutate_if(is.numeric, round, 3) %>%
  select(Parameter, Estimate, "Standard Error" = "Std. Error", "t-value" = "t value", "p-value" = "Pr(>|t|)") %>%
  mutate(Parameter = ifelse(Parameter == "(Intercept)", "intercept", ifelse(Parameter == "gendermale", "male", Parameter)))

```





Table 1 shows descriptive statistics of the study variables. There were no missing values in the data and all participants were included for our final analysis. We found that weight (b = `r coefs[coefs$Parameter == "weight", "Estimate"]`,  `r report_p(coefs[coefs$Parameter == "weight", "p-value"])`) and age (b = `r coefs[coefs$Parameter == "age", "Estimate"]`,  `r report_p(coefs[coefs$Parameter == "age", "p-value"])`) are positively associated with height. Once we know the age and weight of a person, gender is not associated with height. It is possible that gender influences both weight and height and therefore the effect of gender only shows if we do not condition on weight. 

  











